---
title: "Linear Regression"
author: "Milagros Mellizo"
date: "`r Sys.Date()`"
output: html_document
---


```{r}
library('tidyverse')

```

## Data Analysis
Import the data set. Make sure the column names are imported as intended.


```{r}
data <- read_csv('wage2015_subsample_inference.csv')
data
```

Missing Values:

```{r}
data %>% summarise_all(~sum(is.na(.)))
```

Descriptive statistics of the variables (mean, standard deviation, percetiles, etc.)

```{r}
summary(data)
```

* As most of the variables are dummy variables, only wage is interprated based in these descriptive statistics.
* The average wage in the sample is 23.410 dollars per hour.
* The 25% richest earn 27.78 dollars per hour or more, while the 25% poorest earn 13.45 or less.
* The median earning per hour is 19.23 dollars, which is lower than the average wage and shows the asymmetry in the wage distribution of the sample. 


### DF Women : 
Women with a college graduate degree or above have a wage corresponding to the 25% richest

```{r}
data_women_clg_75 <- filter(data, (sex == 1) & (clg==1|ad==1) & (wage > quantile(data$wage,probs = c(0.25,0.5,0.75))[3]))

data_women_clg_75
```

```{r}
print(paste(dim(data_women_clg_75)[1],'women with a college graduate degree (clg) or above have a wage corresponding to the 25% richest of the sample'))
```

### DF Men : 
Men with a high school graduate degree (hsg) or below have a wage corresponding to the 25% richest

```{r}
data_men_hsg_75 <- filter(data, (sex == 0) & (shs==1|hsg==1) & (wage > quantile(data$wage,probs = c(0.25,0.5,0.75))[3]))

data_men_hsg_75
```

```{r}
print(paste(dim(data_men_hsg_75)[1],'men with a high school graduate degree (hsg) or below have a wage corresponding to the 25% richest of the sample'))
```

### DF Log Wage (Y) :
Containing only the log(wage)
```{r}
log_wage <- select(data, lwage)
log_wage
```

Array:
```{r}
y <- data$lwage
```


### DF Features (X) :
Containing every variable of the data set but the wage related variables
```{r}
features <- select(data, -wage, -lwage)
features
```


### Predictors Array:

*Basic Model*

$sex+exp1+hsg+scl+clg+ad+se+we+ne+dummy(occ2)+dummy(ind2)$

```{r}
X_basic <- model.matrix(~ sex + hsg + scl + clg + ad + so + we + ne + exp1 + C(occ2) + C(ind2), data = data)
```

*Flexible Model*

$sex+exp1+exp1+exp2+exp3+exp4+hsg+scl+clg+ad+se+we+ne+dummy(occ2)+dummy(ind2)+(exp1+exp2+exp3+exp4)×(hsg+scl+clg+ad+se+we+ne+dummy(occ2)+dummy(ind2))$

```{r}
X_flexible <- model.matrix(~ sex + (exp1 + exp2 + exp3 + exp4) * (hsg + scl + clg + ad + so + we + ne + C(occ2) + C(ind2)), data = data)
```

*Extra Flexible Model*

$sex+(exp1+exp2+exp3+exp4+hsg+scl+clg+ad+se+we+ne+dummy(occ2)+dummy(ind2))^2$

```{r}
X_extra_flexible <- model.matrix(~ sex + (exp1 + exp2 + exp3 + exp4 + hsg + scl + clg + ad + so + we + ne + C(occ2) + C(ind2)) ^ 2, data = data)
```


## Linear Regression

### Training sample (80% of the data) and a test sample.

```{r}
train_sample <- runif(length(y)) < 0.8
```

```{r}
test_sample <- !train_sample
```

### Estimations

*Basic Model*
```{r}
basic_model <- lm(lwage ~ sex + hsg + scl + clg + ad + so + we + ne + exp1 + C(occ2) + C(ind2), data = data, subset = train_sample)

anova(basic_model)
```

*Flexible Model*
```{r}
flexible_model <- lm(lwage ~ sex + (exp1 + exp2 + exp3 + exp4) * (hsg + scl + clg + ad + so + we + ne + C(occ2) + C(ind2)), data = data, subset = train_sample)

anova(flexible_model)
```

*Extra Flexible Model*
```{r}
extra_flexible_model <- lm(lwage ~ sex + (exp1 + exp2 + exp3 + exp4 + hsg + scl + clg + ad + so + we + ne + C(occ2) + C(ind2)) ^ 2, data = data, subset = train_sample)

anova(extra_flexible_model)
```


## Linear Regression

Split each of the dataframes created (basic, flexible and extra-flexible models) into a training sample (80% of the data) and a test sample.

```{r}
train_sample <- runif(length(y)) < 0.8
```
```{r}
test_sample <- !train_sample
```

Estimates models


*Basic Model*
```{r}
basic_model <- lm(lwage ~ sex + hsg + scl + clg + ad + so + we + ne + exp1 + C(occ2) + C(ind2), data = data, subset = train_sample)
```


*Flexible Model*
```{r}
flexible_model <- lm(lwage ~ sex + (exp1 + exp2 + exp3 + exp4) * (hsg + scl + clg + ad + so + we + ne + C(occ2) + C(ind2)), data = data, subset = train_sample)
```

*Extra Flexible Model*
```{r}
extra_flexible_model <- lm(lwage ~ sex + (exp1 + exp2 + exp3 + exp4 + hsg + scl + clg + ad + so + we + ne + C(occ2) + C(ind2)) ^ 2, data = data, subset = train_sample)
```


### Report 
the MSE and the R2 for both samples of each model, and the adjusted R2 for the training sample


*Basic Model*
```{r}
mse_train_basic <- mean(basic_model$residuals ^ 2)
r2_train_basic <- 1 - mse_train_basic / var(y[train_sample])
adjr2_train_basic <- 1 - length(y) / (length(y) - ncol(X_basic)) * mse_train_basic / var(y)
mse_test_basic <- mean((predict(basic_model, subset(data, subset = test_sample)) - y[test_sample]) ^ 2)
r2_test_basic <- 1 - mse_test_basic / var(y[test_sample])

print(paste("Training MSE for basic model:", mse_train_basic))
print(paste("Training R2 for basic model:", r2_train_basic))
print(paste("Adjusted training R2 for basic model:", adjr2_train_basic))
print(paste("Testing MSE for basic model:", mse_test_basic))
print(paste("Testing R2 for basic model:", r2_test_basic))
```

*Flexible Model*
```{r}
mse_train_flexible <- mean(flexible_model$residuals ^ 2)
r2_train_flexible <- 1 - mse_train_flexible / var(y)
adjr2_train_flexible <- 1 - length(y) / (length(y) - ncol(X_flexible)) * mse_train_flexible / var(y)
mse_test_flexible <- mean((predict(flexible_model, subset(data, subset = test_sample)) - y[test_sample]) ^ 2)
r2_test_flexible <- 1 - mse_test_flexible / var(y[test_sample])

print(paste("Training MSE for flexible model:", mse_train_flexible))
print(paste("Training R2 for flexible model:", r2_train_flexible))
print(paste("Adjusted training R2 for flexible model:", adjr2_train_flexible))
print(paste("Testing MSE for flexible model:", mse_test_flexible))
print(paste("Testing R2 for flexible model:", r2_test_flexible))
```


*Extra Flexible Model*
```{r}
mse_train_extra_flexible <- mean(extra_flexible_model$residuals ^ 2)
r2_train_extra_flexible <- 1 - mse_train_extra_flexible / var(y)
adjr2_train_extra_flexible <- 1 - length(y) / (length(y) - ncol(X_extra_flexible)) * mse_train_extra_flexible / var(y)
mse_test_extra_flexible <- mean((predict(extra_flexible_model, subset(data, subset = test_sample)) - y[test_sample]) ^ 2)
r2_test_extra_flexible <- 1 - mse_test_extra_flexible / var(y[test_sample])

print(paste("Training MSE for flexible model:", mse_train_extra_flexible))
print(paste("Training R2 for flexible model:", r2_train_extra_flexible))
print(paste("Adjusted training R2 for flexible model:", adjr2_train_extra_flexible))
print(paste("Testing MSE for flexible model:", mse_test_extra_flexible))
print(paste("Testing R2 for flexible model:", r2_test_extra_flexible))
```
